---
title: "Summarizing the time series runs"
author: "Liz Hobson"
date: "August 2022"
output: html_document
---

This is Step 2 of the analysis/plotting scripts. NOTE: this script can take a long time to run!!

This script is the initial summary of raw time series data from model runs. Output of this script is summarized data which is saved in the “Cached_data” folder. Two types of summaries are saved:

1: “summary.fixation”… type files, which save data on the generation at which trait variation ceased

2: “summary.preferencesXthresholds”… type files, which save data on the patterns of trait preferences in relation to a series of “thresholds” (We use the 25%/75% data in the main paper, but more are summarized here).


```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(reshape2)
library(scales)
library(stringr)
library(purrr)
library(tidyr)
library(zoo)
library(tidyverse)
library(beepr)

matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}

cv <- function(x) 100*( sd(x)/mean(x))

remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}

#supress warnings
options(warn=-1)
```


#Import data & rbind together (all compiled data from normal runs)
```{r}
data.path <- "Rawdata_compiled/"

#runs.fd.f <- read.csv(paste0(data.path,"compiledruns.fd.f.csv"))  #glimpse(runs.fd.f)
runs.fd.x <- read.csv(paste0(data.path, "compiledruns.fd.x.csv"))  #glimpse(runs.fd.x)
#runs.fde.f <- read.csv(paste0(data.path, "compiledruns.fde.f.csv"))  #glimpse(runs.fde.f)
runs.fde.x <- read.csv(paste0(data.path, "compiledruns.fde.x.csv"))  #glimpse(runs.fde.x)
#runs.ncs.f <- read.csv(paste0(data.path, "compiledruns.ncs.f.csv")) #glimpse(runs.ncs.f)
runs.ncs.x <- read.csv(paste0(data.path, "compiledruns.ncs.x.csv")) #glimpse(runs.ncs.x)
#runs.sc.f <- read.csv(paste0(data.path, "compiledruns.sc.f.csv")) #glimpse(runs.sc.f)
runs.sc.x <- read.csv(paste0(data.path, "compiledruns.sc.x.csv")) #glimpse(runs.sc.x)
runs.fd.fa <- read.csv(paste0(data.path, "compiledruns.fd.fa.csv"))  #glimpse(runs.fd.fa)
runs.fd.fs <- read.csv(paste0(data.path, "compiledruns.fd.fs.csv"))  #glimpse(runs.fd.fs)
runs.fde.xpt1 <- read.csv(paste0(data.path, "compiledruns.fde.xpt1.csv"))  #glimpse(runs.fde.xpt1)
runs.pnov <- read.csv(paste0(data.path, "compiledruns.pnov.x.csv"))  #glimpse(runs.pnov)
runs.pfd.x <- read.csv(paste0(data.path, "compiledruns.pfd.x.csv"))

all.runs <- rbind.data.frame( 
                             runs.fd.x,
                             runs.fde.x,
                             runs.ncs.x,
                             runs.sc.x,
                             runs.fd.fa, 
                             runs.fd.fs,
                             runs.fde.xpt1, 
                             runs.pfd.x,
                             runs.pnov
                             ) 

glimpse(all.runs)


# runs have rowIDs, delete that first column
all.runs <- select(all.runs, -1) #glimpse(all.runs)
#head(all.runs)
```


#separate parts of file names
```{r}
# for col 1 which has full names, split it out then split the parts
all.runs.split.names <- all.runs[,1] #head(all.runs[,1]) #head(all.runs.split.names)

all.runs.split.names <- strsplit(as.character(all.runs.split.names), ".", fixed = TRUE) #head(all.runs.split.names)

#unlist data from one rowxcolumn cell
all.runs.split.names <- as.data.frame(do.call(rbind, all.runs.split.names)) #head(all.runs.split.names) #str(all.runs.split.names)

# add informative column names
colnames(all.runs.split.names) <- c("param", "model", "setting", "group.size", "run") #head(all.runs.split.names)

all.runs.split.names$runID <- sub('.', '', all.runs.split.names$run)
all.runs.split.names$group.size <- as.numeric(sub('.', '', all.runs.split.names$group.size)) 

all.runs.split.names$runID <- as.numeric(all.runs.split.names$runID) #str(all.runs.split.names) unique(all.runs.split.names$setting)


#bind back to all.runs
all.runsWmeta <- cbind.data.frame(all.runs.split.names, all.runs) #str(all.runsWmeta) #str(all.runs)

#add individual run ID name
all.runsWmeta$fileWsa <- paste0(all.runsWmeta$file, "_s", all.runsWmeta$s, "_a", all.runsWmeta$a)

#str(all.runsWmeta)
#test <- head(subset(all.runsWmeta, model=="fd" & setting=="f" & param=="p" & N=="N10" & runID=="3" & s==0 & a==0)) #names(test)

all.runsWmeta <- all.runsWmeta %>% select(-run, -file, -fileWsa) #str(all.runsWmeta) length(all.runsWmeta$runID)
unique(all.runsWmeta$setting)
unique(all.runsWmeta$param)
names(all.runsWmeta)
unique(all.runsWmeta$group.size)

glimpse(all.runsWmeta)
```

# CHECK OUTPUT OF RUNS
Should be 110 subparts per model X params X N X run number (and 2200 per model X params X N)
```{r, echo=TRUE}
run.meta.summary <- all.runsWmeta %>% dplyr::group_by(param, model, setting, group.size) %>% tally() 

#run.meta.summary
unique(run.meta.summary$n) #all of these should have 2200 runs each

```


# LOOP TO SUMMARIZE

## Set up preconditions
```{r}

#filter data to retain only the parameters of interest
all.runsWmeta.main_params <- subset(all.runsWmeta, param=="ta" | param=="p")
unique(all.runsWmeta.main_params$model)
unique(all.runsWmeta.main_params$setting)
unique(all.runsWmeta.main_params$group.size)
unique(all.runsWmeta.main_params$param)

#set unique things to loop over names(all.runsWmeta.sub)
key.MSNsaR <- unique(all.runsWmeta.main_params[c("model", "setting", "group.size", "runID", "s", "a")]) #unique(df[c("yad", "per")])

# SORT
key.MSNsaR <- key.MSNsaR %>% arrange(model, setting, group.size, s, a, runID)
nrow(key.MSNsaR)
unique(key.MSNsaR$model)

# set up thresholds 
lower <- c(0.10, 0.20, 0.25, 0.30, 0.40)
upper <- c(0.60, 0.70, 0.75, 0.80, 0.90)

threshold.pairs <- as.data.frame(cbind(lower, rev(upper)))
colnames(threshold.pairs) <- c("lower", "upper")
threshold.pairs$lower.lab <- c(10, 20, 25, 30, 40)
threshold.pairs$upper.lab <- c(90,80,75,70,60)

```


# SUMMARIZE IN LOOP 

rle() finds the run length for repeated values more extreme than thresholds. "Run" (number repeated in a row) must be >1 (at least two values in a row more extreme than thresholds)

Remove outliers: https://stackoverflow.com/questions/4787332/how-to-remove-outliers-from-a-dataset/4788102#4788102
```{r}
# get start time to find run time at end of loop
start.time <- Sys.time()

#SUBSET TO RUN JUST PARTICULAR MODEL/SETTING COMBO
#key.MSNsaR <- subset(key.MSNsaR, model=="fd" & setting=="x")  #glimpse(key.MSNsaR)

key.models <- as.character(unique(all.runsWmeta.main_params$model))
unique.loop.MSNkeys <- unique(all.runsWmeta.main_params[c("model", "setting", "group.size")]) 


#to run just one model variant, adjust here:
key.models <- "pfd"
unique.loop.MSNkeys <- subset(unique.loop.MSNkeys, model=="pfd")

#LOOP
# for each model
for(MODEL in 1:length(key.models)) { #MODEL=1
  loop.model <- key.models[MODEL]
  #loop.data.4keys <- subset(all.runsWmeta.main_params, model==loop.model) 
  loop.MSN.keysXmodel <- subset(unique.loop.MSNkeys, model==loop.model)
  key.setting <- as.character(unique(loop.MSN.keysXmodel$setting)) #list settings in focal model

  for(SETTING in 1:length(key.setting)) { #SETTING=1 
    loop.setting <- key.setting[SETTING] 
    loop.MSNkeysXsetting <- subset(loop.MSN.keysXmodel, setting==loop.setting)

    key.groupsizes <- as.numeric(unique(loop.MSNkeysXsetting$group.size))
  

    for(GROUPN in 1:length(key.groupsizes)) { #GROUPN=1
      loop.groupn <- key.groupsizes[GROUPN]
      
      print(paste("current model:", loop.model, loop.setting, "group size", loop.groupn, sep=" "))

      # filter keys
      loop.MSNRsa <- key.MSNsaR %>% filter(model==loop.model, setting==loop.setting, group.size==loop.groupn) #glimpse(loop.MSNRsa)
      
      
      # Initiate blank dataframes (write to these once per model/setting/group combo)
      summary.fixation <- data.frame(
                         model=character(),
                         stetting=character(),
                         group.size=numeric(),
                         runID=numeric(),
                         s=numeric(),
                         a=numeric(),
                         param=character(),
                         first.stable.gen=numeric(), 
                         n.gens.stable=numeric(), 
                         first.stable.gen.fuzzy=numeric(), 
                         n.gens.stable.fuzzy=numeric()
                         )

      
        summary.thresholds <- data.frame(
                                   model=character(),
                                   stetting=character(),
                                   groupn=numeric(),
                                   run=numeric(),
                                   s=numeric(),
                                   a=numeric(),
                                   param=character(),
                                   thresholds=character(),
                                   n_lower=numeric(),
                                   n_upper=numeric(),
                                   n.extreme.shifts=numeric(),
                                   #all streaks (potentially including any outliers)
                                   ext.streak.median=numeric(),
                                   ext.streak_3gen.n=numeric(),
                                   ext.streak_5gen.n=numeric(),
                                   ext.streak_10gen.n=numeric(),
                                   ext.streak_3gen.median=numeric(),
                                   ext.streak_5gen.median=numeric(),
                                   ext.streak_10gen.median=numeric(),
                                   #without outliers
                                   outliers.n=numeric(),
                                   ext.streak.median_no.outliers=numeric(),
                                   ext.streak_3gen.n.noouts=numeric(),
                                   ext.streak_5gen.n.noouts=numeric(),
                                   ext.streak_10gen.n.noouts=numeric(),
                                   #number streaks of certain length in upper & lower individually
                                   ext.streak_3gen.n_Up.noouts=numeric(),
                                   ext.streak_5gen.n_Up.noouts=numeric(),
                                   ext.streak_10gen.n_Up.noouts=numeric(),
                                   ext.streak_3gen.n_Lw.noouts=numeric(),
                                   ext.streak_5gen.n_Lw.noouts=numeric(),
                                   ext.streak_10gen.n_Lw.noouts=numeric(),
                                   #summary of streak length
                                   ext.streak_3gen.median.noouts=numeric(),
                                   ext.streak_5gen.median.noouts=numeric(),
                                   ext.streak_10gen.median.noouts=numeric()
                                   )
      
      # Initiate progress bar
      n.loops<-nrow(loop.MSNRsa) 
      pb <- txtProgressBar(min = 0, max = n.loops, style = 3)
      

      for(KEY in 1:nrow(loop.MSNRsa)) { #KEY=600
        
        # Update the progress bar
        setTxtProgressBar(pb, KEY)
        
        # extract temp values for loop
        loop.run <- loop.MSNRsa[KEY,4]
        loop.s <- loop.MSNRsa[KEY,5]
        loop.a <- loop.MSNRsa[KEY,6]
        
        loop.meta <- cbind.data.frame(loop.model, loop.setting, loop.groupn, loop.run, loop.s, loop.a)
        colnames(loop.meta) <- c("model", "setting", "groupn", "run", "s", "a")
        
        # extract data for this particular MSNRsa key
        loop.data.ta <- all.runsWmeta.main_params %>% filter(model==loop.model, 
                                                     setting==loop.setting, 
                                                     group.size==loop.groupn, 
                                                     runID==loop.run,
                                                     s==loop.s,
                                                     a==loop.a,
                                                     param=="ta")
        
        loop.data.p <- all.runsWmeta.main_params %>% filter(model==loop.model, 
                                                     setting==loop.setting, 
                                                     group.size==loop.groupn, 
                                                     runID==loop.run,
                                                     s==loop.s,
                                                     a==loop.a,
                                                     param=="p")        

        loop.data.ta.gendata <- loop.data.ta[1,8:ncol(loop.data.ta)]
        loop.data.ta.gendata.melt <- reshape2::melt(loop.data.ta.gendata, id.vars = NULL)  
        loop.data.ta.gendata.melt <- loop.data.ta.gendata.melt %>% arrange(variable) #make sure data are sorted by generation
        
        loop.data.p.gendata <- loop.data.p[1,8:ncol(loop.data.p)]
        loop.data.p.gendata.melt <- reshape2::melt(loop.data.p.gendata, id.vars = NULL)  
        loop.data.p.gendata.melt <- loop.data.p.gendata.melt %>% arrange(variable) #make sure data are sorted by generation
        
        #combine ta and p data into single dataframe
        loop.data.ta_p <- cbind.data.frame(loop.data.ta.gendata.melt[,1:2], loop.data.p.gendata.melt[,2])
        colnames(loop.data.ta_p) <- c("gen", "ta", "p")  
  
        #make generation column numeric
        loop.data.ta_p$gen <- as.numeric(sub('.', '', loop.data.ta_p$gen)) #head(loop.data.ta_p) 

        #replace any NAs with 0
        loop.data.ta_p[is.na(loop.data.ta_p)] <- 0

        #make sure data are sorted by generationsort
        FWD_data <- loop.data.ta_p %>% arrange(gen) #head(FWD_data.ta)
        
        #reverse sort (to find stabilization)
        REV_data <- loop.data.ta_p %>% arrange(-gen)

  
    ########## Summarize Trait A data for trait fixation
        
        #stabilization: at what generation does raw ts stabilize? 
        REV_data$lead.ta <- lead(REV_data$ta)
        REV_data[is.na(REV_data)] <- 0 #make all rows that are na 0's (gen 100=ta 0)
        REV_data$lead.diff <- REV_data$lead.ta-REV_data$ta
        REV_data[100,5] <- 0 #make last row lead difference 0
        
        first.non0 <- REV_data %>%
                          filter(lead.diff > 0) %>%
                          slice(1) %>%
                          pull(gen)
        
        first.non000001 <- REV_data %>%
                          filter(abs(lead.diff) > 0.000001) %>%
                          slice(1) %>%
                          pull(gen)
        
        #if the result from above is nothing, then the system was always stable
          #return generation 1 at which stability occurred
          #otherwise, return the generation at which the above is first satisfied (first stable gen)
          #if system NEVER stabilized, return 100 for first stable gen and 0 for n gens stable
        first.stable.gen <- ifelse(length(first.non0)==0, 1, first.non0)
        first.stable.gen.fuzzy <- ifelse(length(first.non000001)==0, 1, first.non000001)
       
        n.gens.stable <- 100-first.stable.gen
        n.gens.stable.fuzzy <- 100-first.stable.gen.fuzzy
        
        FWD_data.fuzzy <- subset(FWD_data, gen<=first.stable.gen.fuzzy)

        ### COMPILE SUMMARY STATS for Trait A
        compiled.fixation <- cbind.data.frame(
                        loop.meta,
                        first.stable.gen, n.gens.stable, 
                        first.stable.gen.fuzzy, n.gens.stable.fuzzy
                        )
        
        #bind compiled to main summary stats
        summary.fixation <- rbind.data.frame(summary.fixation, compiled.fixation) 
        
        # Quantify streaks & extreme shifts for both trait A and preference for each threshold
        for(th in 1:nrow(threshold.pairs)){ #th=3
          
           #for loop, name lower and upper thresholds (for labeling data later)
          thrsh.lower <- threshold.pairs$lower.lab[th]
          thrsh.upper <- threshold.pairs$upper.lab[th]
          loop.thresholds <- paste(thrsh.lower, thrsh.upper, sep="_")

          #for(PARAM in 1:length(key.param)) { #PARAM=1 -- use FWD_data.fuzzy[,PARAM+1] for muliple params (ta first)
            
          PARAM <- "p"
          
          #count the number of values more extreme than thresholds
          n_lower <- sum(FWD_data.fuzzy$p<threshold.pairs$lower[th])
          n_upper <- sum(FWD_data.fuzzy$p>threshold.pairs$upper[th])
          
          ### Mark when values are more extreme than thresholds (use to find runs/streaks where values more extreme)
          FWD_data.fuzzy$YN_lower <- ifelse(FWD_data.fuzzy$p<threshold.pairs$lower[th], "Lw", "n")
          FWD_data.fuzzy$YN_upper <- ifelse(FWD_data.fuzzy$p>threshold.pairs$upper[th], "Up", "n")
          FWD_data.fuzzy$YN_both <- ifelse(FWD_data.fuzzy$YN_lower=="Lw", "Lw", ifelse(FWD_data.fuzzy$YN_upper=="Up", "Up", "n"))
          
          #subset to include streak data only for extreme values (exclude streaks in non-extreme zone)
          YN_extreme <- FWD_data.fuzzy$YN_both[FWD_data.fuzzy$YN_both!="n"] 
          
          #make a version that is numeric 
          YN_extreme.num <- YN_extreme
          YN_extreme.num[YN_extreme.num=="Up"] <- 1
          YN_extreme.num[YN_extreme.num=="Lw"] <- -1
          YN_extreme.num <- as.numeric(YN_extreme.num)
          
          #remove duplicated values from the vector, just save "shifts" between non-extreme, high-extreme, low-extreme 
          # if there aren't any shifts (YN_extreme.num=NA) this will just return NA
          YN_extreme.shifts <- YN_extreme.num[YN_extreme.num != lag(YN_extreme.num, default = !YN_extreme.num[1])]
          
          #how many shifts between extremes were there? 
          # if only in one extreme, length will be 1; one extreme then other=2, cycle from one to other back to first=3
          # subtract 1 from total in case there's only a 1 or -1 in YN_extreme.shifts
          # if there aren't any shifts, length will be 1 (for the NA) and below will return n.extreme.shifts=0
          n.extreme.shifts <- ifelse(length(YN_extreme.shifts)>1, length(YN_extreme.shifts)-1, 0)
          
          #find where values exceed thresholds several generations in a row (a "streak") using only extreme values
          rle.streaks <- rle(FWD_data.fuzzy$YN_both)
          
          # check for whether there is any data, if none, return 0's/NONE 
          # is.logical is FALSE when there's data, TRUE when no data
          ifelse(is.logical(rle.streaks[1])==TRUE, lengths <- 0, lengths <- rle.streaks$lengths)
          ifelse(is.logical(rle.streaks[2])==TRUE, values <- "NONE", values <- rle.streaks$values)
          
          streaks <- cbind.data.frame(lengths, values)

          ext.streaks <- subset(streaks, values!="n" & lengths>1) #CHANGE BACK TO 1
          
          #if there aren't any streaks longer than 2 (two same-category in a row), add NONE to dataframe so it's not empty
          options(stringsAsFactors=FALSE)
          if(nrow(ext.streaks)==0) ext.streaks <- rbind.data.frame(ext.streaks, 
                                                                   cbind.data.frame(lengths=c(NA), values=c(as.character("NONE")))) 
          
            
          # NO POINTS REMOVED (POTENTIALLY WITH OUTLIERS)
          # CALCULATE THESE VERSIONS FOR REFERENCE, BUT USE VALUES WITHOUT OUTLIERS FOR ANALYSES
          ifelse(ext.streaks$values=="NONE", 
                 ext.streak.median <- NA, 
                 ext.streak.median <- median(ext.streaks$lengths))
          
          # return any streak greater than or equal to the desired number of generations
          ifelse(length(ext.streaks$lengths[ext.streaks$lengths>=3])>0, 
                 ext.streak_3gen <- ext.streaks$lengths[ext.streaks$lengths>=3], 
                 ext.streak_3gen <-NA)
          
          ifelse(length(ext.streaks$lengths[ext.streaks$lengths>=5])>0, 
                 ext.streak_5gen <- ext.streaks$lengths[ext.streaks$lengths>=5], 
                 ext.streak_5gen <-NA)
          
          ifelse(length(ext.streaks$lengths[ext.streaks$lengths>=10])>0, 
                 ext.streak_10gen <- ext.streaks$lengths[ext.streaks$lengths>=10], 
                 ext.streak_10gen <-NA)
          
          
          ifelse(is.na(ext.streak_3gen), 
                 ext.streak_3gen.n <- NA, 
                 ext.streak_3gen.n <- length(ext.streak_3gen))
          
          ifelse(is.na(ext.streak_5gen), 
                 ext.streak_5gen.n <- NA, 
                 ext.streak_5gen.n <- length(ext.streak_5gen))
          
         ifelse(is.na(ext.streak_10gen), 
                 ext.streak_10gen.n <- NA, 
                 ext.streak_10gen.n <- length(ext.streak_10gen))
           

         ifelse(is.na(ext.streak_3gen), 
                 ext.streak_3gen.n_Up <- NA, 
                 ext.streak_3gen.n_Up <- length(ext.streaks$values[ext.streaks$lengths>=3 & ext.streaks$values=="Up"]))
          
         ifelse(is.na(ext.streak_5gen), 
                 ext.streak_5gen.n_Up <- NA, 
                 ext.streak_5gen.n_Up <- length(ext.streaks$values[ext.streaks$lengths>=5 & ext.streaks$values=="Up"]))
         
         ifelse(is.na(ext.streak_10gen), 
                 ext.streak_10gen.n_Up <- NA, 
                 ext.streak_10gen.n_Up <- length(ext.streaks$values[ext.streaks$lengths>=10 & ext.streaks$values=="Up"])) 
         
         ifelse(is.na(ext.streak_3gen), 
                 ext.streak_3gen.n_Lw <- NA, 
                 ext.streak_3gen.n_Lw <- length(ext.streaks$values[ext.streaks$lengths>=3 & ext.streaks$values=="Lw"]))
          
        ifelse(is.na(ext.streak_5gen), 
                 ext.streak_5gen.n_Lw <- NA, 
                 ext.streak_5gen.n_Lw <- length(ext.streaks$values[ext.streaks$lengths>=5 & ext.streaks$values=="Lw"]))
         
        ifelse(is.na(ext.streak_10gen), 
                 ext.streak_10gen.n_Lw <- NA, 
                 ext.streak_10gen.n_Lw <- length(ext.streaks$values[ext.streaks$lengths>=10 & ext.streaks$values=="Lw"]))         
          
        ifelse(is.na(ext.streak_3gen), 
                 ext.streak_3gen.median <- NA, 
                 ext.streak_3gen.median <- median(ext.streak_3gen, na.rm=TRUE))
        
        ifelse(is.na(ext.streak_5gen), 
                 ext.streak_5gen.median <- NA, 
                 ext.streak_5gen.median <- median(ext.streak_5gen, na.rm=TRUE))
          
        ifelse(is.na(ext.streak_10gen), 
                 ext.streak_10gen.median <- NA, 
                 ext.streak_10gen.median <- median(ext.streak_10gen, na.rm=TRUE))
        


          # WITHOUT OUTLIERS (OUTLIER STREAK LENGTHS REMOVED)
          # USE THIS VERSION FOR SUMMARIES
          # ** if ext.streak.median==NA then there are no streaks in timeseries
          ifelse(ext.streaks$values=="NONE", 
                 na_outliers <- "NONE", 
                 na_outliers <- remove_outliers(ext.streaks$lengths)) 
          
          ext.streaks.noouts <- cbind.data.frame(ext.streaks, na_outliers)
          no_outliers <- subset(ext.streaks.noouts, !is.na(na_outliers))

          #count of number of outliers (number NAs from remove_outliers)
          ifelse(ext.streaks.noouts$na_outliers=="NONE", 
                 outliers.n <- 0, 
                 outliers.n <- sum(is.na(ext.streaks.noouts$na_outliers)))
          
          #find median streak length without outliers
          ifelse(no_outliers$na_outliers=="NONE",  
                 ext.streak.median_no.outliers <- NA, 
                 ext.streak.median_no.outliers <- median(no_outliers$na_outliers, na.rm=TRUE))

          # return any streak greater than or equal to the desired number of generations
          # assign these within the ifelse statement to retain the vectors
          # is.logical is FALSE when there's data, TRUE when no data
          ifelse(no_outliers$na_outliers=="NONE", 
                 ext.streak_3gen.noouts <- NA, 
              ifelse(no_outliers$na_outliers!="NONE" & nrow(subset(no_outliers, na_outliers>=3))==0,
                     ext.streak_3gen.noouts <- NA,
                     ext.streak_3gen.noouts <- no_outliers$na_outliers[no_outliers$na_outliers>=3]))

          ifelse(no_outliers$na_outliers=="NONE", 
                 ext.streak_5gen.noouts <- NA, 
              ifelse(no_outliers$na_outliers!="NONE" & nrow(subset(no_outliers, na_outliers>=5))==0,
                 ext.streak_5gen.noouts <- NA,
                 ext.streak_5gen.noouts <- no_outliers$na_outliers[no_outliers$na_outliers>=5]))
          
          ifelse(no_outliers$na_outliers=="NONE", 
                 ext.streak_10gen.noouts <- NA, 
              ifelse(no_outliers$na_outliers!="NONE" & nrow(subset(no_outliers, na_outliers>=10))==0,
                 ext.streak_10gen.noouts <- NA,
                 ext.streak_10gen.noouts <- no_outliers$na_outliers[no_outliers$na_outliers>=10]))
         
          # find the number of streaks (compare this to the number of extreme shifts to see if streaks are at both extremes)
          ifelse(is.na(ext.streak_3gen.noouts), 
                 ext.streak_3gen.n.noouts <- NA, 
                 ext.streak_3gen.n.noouts <- length(ext.streak_3gen.noouts))
          
          ifelse(is.na(ext.streak_5gen.noouts), 
                 ext.streak_5gen.n.noouts <- NA, 
                 ext.streak_5gen.n.noouts <- length(ext.streak_5gen.noouts))
          
          ifelse(is.na(ext.streak_10gen.noouts), 
                 ext.streak_10gen.n.noouts <- NA, 
                 ext.streak_10gen.n.noouts <- length(ext.streak_10gen.noouts))

          ifelse(is.na(ext.streak_3gen.noouts), 
                 ext.streak_3gen.n_Up.noouts <- NA, 
                 ext.streak_3gen.n_Up.noouts <- length(ext.streaks.noouts$values[ext.streaks.noouts$lengths>=3 & ext.streaks.noouts$values=="Up"]))
          
         ifelse(is.na(ext.streak_5gen.noouts), 
                ext.streak_5gen.n_Up.noouts <- NA, 
                ext.streak_5gen.n_Up.noouts <- length(ext.streaks.noouts$values[ext.streaks.noouts$lengths>=5 & ext.streaks.noouts$values=="Up"]))
         
         ifelse(is.na(ext.streak_10gen.noouts), 
                ext.streak_10gen.n_Up.noouts <- NA, 
                ext.streak_10gen.n_Up.noouts <- length(ext.streaks.noouts$values[ext.streaks.noouts$lengths>=10 & ext.streaks.noouts$values=="Up"])) 
         
          ifelse(is.na(ext.streak_3gen.noouts),
                 ext.streak_3gen.n_Lw.noouts <- NA, 
                 ext.streak_3gen.n_Lw.noouts <- length(ext.streaks.noouts$values[ext.streaks.noouts$lengths>=3 & ext.streaks.noouts$values=="Lw"]))
          
         ifelse(is.na(ext.streak_5gen.noouts), 
                ext.streak_5gen.n_Lw.noouts <- NA, 
                ext.streak_5gen.n_Lw.noouts <- length(ext.streaks.noouts$values[ext.streaks.noouts$lengths>=5 & ext.streaks.noouts$values=="Lw"]))
         
         ifelse(is.na(ext.streak_10gen.noouts), 
                ext.streak_10gen.n_Lw.noouts <- NA, 
                ext.streak_10gen.n_Lw.noouts <- length(ext.streaks.noouts$values[ext.streaks.noouts$lengths>=10 & ext.streaks.noouts$values=="Lw"])) 
         
          # find median streak length
          ext.streak_3gen.median.noouts <- median(ext.streak_3gen.noouts, na.rm=TRUE)
          ext.streak_5gen.median.noouts <- median(ext.streak_5gen.noouts, na.rm=TRUE)
          ext.streak_10gen.median.noouts <- median(ext.streak_10gen.noouts, na.rm=TRUE)
          
          thresh.loop.compiled <- cbind.data.frame(
                                         loop.meta,
                                         PARAM,
                                         loop.thresholds,
                                         n_lower,
                                         n_upper,
                                         n.extreme.shifts,
                                         #all streaks (potentially including any outliers)
                                         ext.streak.median,
                                         ext.streak_3gen.n,
                                         ext.streak_5gen.n,
                                         ext.streak_10gen.n,
                                         ext.streak_3gen.median,
                                         ext.streak_5gen.median,
                                         ext.streak_10gen.median,
                                         #without outliers
                                         outliers.n,
                                         ext.streak.median_no.outliers,
                                         ext.streak_3gen.n.noouts,
                                         ext.streak_5gen.n.noouts,
                                         ext.streak_10gen.n.noouts,
                                         #number streaks of certain length in upper & lower individually
                                         ext.streak_3gen.n_Up.noouts,
                                         ext.streak_5gen.n_Up.noouts,
                                         ext.streak_10gen.n_Up.noouts,
                                         ext.streak_3gen.n_Lw.noouts,
                                         ext.streak_5gen.n_Lw.noouts,
                                         ext.streak_10gen.n_Lw.noouts,
                                         #summary of all streak lengths
                                         ext.streak_3gen.median.noouts,
                                         ext.streak_5gen.median.noouts,
                                         ext.streak_10gen.median.noouts
                                         )
          
          summary.thresholds <- rbind.data.frame(summary.thresholds, thresh.loop.compiled) 
          #} #end parameter loop
          
          } #end threshold loop
          
        } #end key loop (by runID)
        
         # at end of each group size loop, write data to file and clear garbage
         write_csv(summary.fixation, paste0("Cached_data/summary.fixationXta_", 
                                           paste(loop.meta$model, loop.meta$setting, loop.meta$groupn, sep="_"), 
                                           ".csv")
                  )
      
         write_csv(summary.thresholds, paste0("Cached_data/summary.preferencesXthresholds_", 
                                           paste(loop.meta$model, loop.meta$setting, loop.meta$groupn, sep="_"), 
                                           ".csv")
                  )
          
      } #end group size loop

      # make a "fanfare" sound at end of this loop   
      beepr::beep(sound = 3, expr = NULL)
  
        #clear garbage
        gc()

  } #end setting loop
        
} #end model loop    
       
#make a ping sound if there is an error
beepr::beep_on_error(expr, sound = 1)        


# get end time to find run time at end of loop
end.time <- Sys.time()
run.time <- end.time - start.time 
print(run.time) #2.05 hours run time for 2022-01-02 results
```
