---
title: "Step04b_summarize tseries preferences & traits for heatmaps"
author: "Liz Hobson"
date: "2022-10-02"
output: html_document
---

This script is Step 4b of the analysis/plotting scripts. 

Step 04b is a slight variation of the Step04a script. This variation plots the same time series, but for all Run 02 exemplars instead of all Run 01 exemplars. We plotted the Run 02 exemplars because in one case, the Run 01 exemplar happened to have an unusual pattern that differed from the general pattern seen in the rest of the 20 replicate runs. In this case, we subbed in a Run 02 exemplar to better illustrate the main pattern we highlighted (Figure 4a, the panel for s = 0.1 and a = 0.7 is Run 02; all other time series exemplars are from Run 01s / Step 04a script output).   


# Import data & load packages
```{r}
library(dplyr)
library(tidyverse)
library(reshape2)
library(gplots)
library(RColorBrewer)
library(extraoperators)
library(stringr)          

cv <- function(x) 100*( sd(x)/mean(x))

# set data path
compiled.data.path <- "Cached_data/"

file.list <- as.data.frame(list.files(compiled.data.path, recursive=TRUE))
colnames(file.list) <- "file.name"

# extract just file names with fixation
file.list_trait.fixation <- as.data.frame(file.list[stringr::str_detect(file.list$file.name, "fixation"), ])  # Extract matching rows with str_detect
colnames(file.list_trait.fixation) <- "file.name"
#compile trait data
data_traits <- do.call(rbind, lapply(paste0(compiled.data.path,file.list_trait.fixation$file.name), function(x) read.csv(x, stringsAsFactors = FALSE)))


# extract just file names with preference
file.list_preferences <- as.data.frame(file.list[stringr::str_detect(file.list$file.name, "preference"), ]) # Extract matching rows with str_detect
colnames(file.list_preferences) <- "file.name"
#compile preference data
data_preferences <- do.call(rbind, lapply(paste0(compiled.data.path,file.list_preferences$file.name), function(x) read.csv(x, stringsAsFactors = FALSE)))

```

#checks
```{r}

CHECK.streak_p.pnov_test <- subset(data_preferences, model=="pnov")
glimpse(CHECK.streak_p.pnov_test)
unique(CHECK.streak_p.pnov_test$n.extreme.shifts)


glimpse(data_traits)
traits.summary <- data_traits %>% group_by(model, setting, groupn,s,a) %>% summarize(nruns=n())
unique(traits.summary$nruns)

```

# Data wrangling
```{r}
# sort
data_traits <- data_traits %>% arrange(model, setting, groupn,s,a, run) #names(data_traits)
data_preferences <- data_preferences %>% arrange(model, setting, groupn,s,a, run, loop.thresholds) 
```

# Summarize trait data for Trait A
Find the median generation at which fixation occurred for all runs where fixation did occur (exclude any runs where variation was maintained)
```{r}
unique(data_traits$first.stable.gen.fuzzy)

compiled.ta <- data_traits %>% 
                  dplyr::group_by(model, setting, groupn, s, a) %>% 
                  dplyr::summarize(
                      # for strict stabilization definition
                        ta.stabilized_Nruns=sum(first.stable.gen<100), 
                        ta.first.stabilized.gen_median=median(first.stable.gen[first.stable.gen<100]),
                      # for fuzzy stabilization definition
                        ta.runs.stabilized.fuzzy_N=sum(first.stable.gen.fuzzy<100), #how many runs n.gens.stable>0 (NA if it didn't stabilize)
                        ta.first.stabilized.gen.fuzzy_median=median(first.stable.gen.fuzzy[first.stable.gen.fuzzy<100]),
                      ta.nruns.variation.maintained_fuzzy=20-ta.runs.stabilized.fuzzy_N
                )

#glimpse(compiled.ta)
unique(compiled.ta$ta.nruns.variation.maintained_fuzzy)

```

#Import raw timeseries data & rbind together
```{r}
# set data path
raw.data.path <- "Rawdata_compiled/"

# import data
runs.fd.fa <- read.csv(paste0(raw.data.path, "compiledruns.fd.fa.csv"))  
runs.fd.fs <- read.csv(paste0(raw.data.path, "compiledruns.fd.fs.csv")) 
runs.fd.x <- read.csv(paste0(raw.data.path, "compiledruns.fd.x.csv")) 
runs.fde.x <- read.csv(paste0(raw.data.path, "compiledruns.fde.x.csv")) 
runs.fde.xpt1 <- read.csv(paste0(raw.data.path, "compiledruns.fde.xpt1.csv"))  
runs.ncs.x <- read.csv(paste0(raw.data.path, "compiledruns.ncs.x.csv")) 
runs.pfd.x <- read.csv(paste0(raw.data.path, "compiledruns.pfd.x.csv")) 
runs.pnov.x <- read.csv(paste0(raw.data.path, "compiledruns.pnov.x.csv"))  
runs.sc.x <- read.csv(paste0(raw.data.path, "compiledruns.sc.x.csv")) 

# compile data
all.runs <- rbind.data.frame(runs.fd.fa, 
                             runs.fd.fs,
                             runs.fd.x, 
                             runs.fde.x,
                             runs.fde.xpt1,
                             runs.ncs.x,
                             runs.pfd.x,
                             runs.pnov.x,
                             runs.sc.x
                             ) 
# check compiled data
glimpse(all.runs)

# runs have rowIDs, delete that first column
all.runs <- select(all.runs, -1) #glimpse(all.runs)
head(all.runs)
```

## Extract metadata from raw timeseries files
```{r}
#separate parts of file names
# for col 1 which has full names, split it out then split the parts
all.runs.split.names <- all.runs[,1] #head(all.runs[,1]) #head(all.runs.split.names)

all.runs.split.names <- strsplit(as.character(all.runs.split.names), ".", fixed = TRUE) #head(all.runs.split.names)

#unlist data from one rowxcolumn cell
all.runs.split.names <- as.data.frame(do.call(rbind, all.runs.split.names)) #head(all.runs.split.names) #str(all.runs.split.names)

# add informative column names
colnames(all.runs.split.names) <- c("param", "model", "setting", "group.size", "run") #head(all.runs.split.names)

all.runs.split.names$runID <- sub('.', '', all.runs.split.names$run)
all.runs.split.names$group.size <- as.numeric(sub('.', '', all.runs.split.names$group.size)) 

all.runs.split.names$runID <- as.numeric(all.runs.split.names$runID) #str(all.runs.split.names) unique(all.runs.split.names$setting)


#bind back to all.runs
all.runsWmeta <- cbind.data.frame(all.runs.split.names, all.runs) #str(all.runsWmeta) #str(all.runs)

#add individual run ID name
all.runsWmeta$fileWsa <- paste0(all.runsWmeta$file, "_s", all.runsWmeta$s, "_a", all.runsWmeta$a)

#str(all.runsWmeta)
#test <- head(subset(all.runsWmeta, model=="fd" & setting=="f" & param=="p" & N=="N10" & runID=="3" & s==0 & a==0)) #names(test)

all.runsWmeta <- all.runsWmeta %>% select(-run, -file, -fileWsa) #str(all.runsWmeta) length(all.runsWmeta$runID)
unique(all.runsWmeta$setting)
unique(all.runsWmeta$param)
names(all.runsWmeta)
unique(all.runsWmeta$group.size)

glimpse(all.runsWmeta)
head(all.runsWmeta)
```

### CHECK OUTPUT OF RUNS
Should be 110 subparts per model X params X N X run number (and 2200 per model X params X N)
```{r, echo=TRUE}
run.meta.summary <- all.runsWmeta %>% dplyr::group_by(param, model, setting, group.size) %>% tally() 

#run.meta.summary
unique(run.meta.summary$n) #all of these should have 2200 runs each

```


# Import shortcut data ** CHANGED RUN ID HERE TO RUN02
```{r}
all.runsWmeta.params_ta.tb.p <- subset(all.runsWmeta, param=="ta" | param=="tb" | param=="p")
example.ts <- subset(all.runsWmeta.params_ta.tb.p, runID==2 & group.size==30)
example.ts <- subset(example.ts, s==0.1 | s==0.4)
example.ts <- subset(example.ts, a==2 | a==7)
names(example.ts)
example.ts.melt <- reshape2::melt(example.ts, id.vars = c("model", "setting", "group.size", "param", "runID", "s", "a"))
head(example.ts.melt)
#rename measure column
example.ts.melt$gen <- as.numeric(sub('.', '', example.ts.melt$variable))
example.ts.melt <- example.ts.melt %>% arrange(model, setting, group.size, param, s, a, runID, gen)
#add MS key for easier plotting
example.ts.melt$key.MS <- paste(example.ts.melt$model, 
                                example.ts.melt$setting, 
                                sep="."
                                )


# Retain just the summary data for run 1
run02s_trait.summ <- subset(data_traits, run=="2" & groupn==30)
run02s_trait.summ.sa <- subset(run02s_trait.summ, s==0.2 | s==0.8)
run02s_trait.summ.sa <- subset(run02s_trait.summ.sa, a==2 | a==7)
#add MS key for easier plotting
run02s_trait.summ.sa$key.MS <- paste(run02s_trait.summ.sa$model, 
                                run02s_trait.summ.sa$setting, 
                                sep="."
                                )
names(run02s_trait.summ.sa) #use first.stable.gen.fuzzy to plot

```

# Summarize preference data 
```{r}

#str(streak_p)
compiled.preferences <- data_preferences %>% dplyr::group_by(model, setting, groupn, s, a, loop.thresholds) %>% 
  dplyr::summarize(
    # median number of values in lower/upper thresholds
    n_lower.median=median(n_lower, na.rm=TRUE), 
    n_upper.median=median(n_upper, na.rm=TRUE), 
    
    # number extreme shifts (1=went from one extreme to other once, no NA, >=2 at least one cycle)
    extcircuit.nreps.PRESENT=sum(n.extreme.shifts>=2), #count n replicates with shifts >=2 (at least one "cycle")
    nshiftsXrep.MEDIAN=median(n.extreme.shifts), #median across all replicates, INCLUDING 0
    nshiftsXrep.MEDIAN.no0=median(n.extreme.shifts[n.extreme.shifts>=2]), #median across all replicates, EXCLUDING 0
    
    # # count number runs with outliers (0-n, no NAs)
    # outliers.n.PRESENT=sum(outliers.n>0), #n replicates with any outliers (count where >0)
    # outliers.n.MEDIAN=median(outliers.n), #median number of outliers per replicate, including 0
    
    # ALL DATA, INCLUDING OUTINER STREAK LENGTHS
    # count of replicates with extreme streaks present (NA-n, NAs present)
    gen3.nreps.streak.Wos.PRESENT=sum(!is.na(ext.streak_3gen.n)), #count n replicates with any streaks more than 3 gen (where not NA) 
    gen5.nreps.streak.Wos.PRESENT=sum(!is.na(ext.streak_5gen.n)), #count n replicates with any streaks more than 5 gen (where not NA) 
    gen10.nreps.streak.Wos.PRESENT=sum(!is.na(ext.streak_10gen.n)), #count n replicates with any streaks more than 10 gen (where not NA) 
    
    # mean number of streaks per replicate (NA-n, NAs present)
    gen3.NstreakXrep.Wos.MEAN=mean(ext.streak_3gen.n, na.rm=TRUE),#mean of number of streaks / replicate, na.rm=TRUE (streaks more than 3 gen)
    gen5.NstreakXrep.Wos.MEAN=mean(ext.streak_5gen.n, na.rm=TRUE), #mean of number of streaks / replicate, na.rm=TRUE (streaks more than 5 gen)
    gen10.NstreakXrep.Wos.MEAN=mean(ext.streak_10gen.n, na.rm=TRUE), #mean of number of streaks / replicate, na.rm=TRUE (streaks more than 10 gen)
    
    # mean of median streak length per replicate (NA-n, NAs present)  
    gen3.medianstreaklenXrep.Wos.MEAN=mean(ext.streak_3gen.median, na.rm=TRUE), #mean of median length of streaks +3gen / replicate, na.rm=TRUE 
    gen5.medianstreaklenXrep.Wos.MEAN=mean(ext.streak_5gen.median, na.rm=TRUE), #mean of median length of streaks +5gen / replicate, na.rm=TRUE 
    gen10.medianstreaklenXrep.Wos.MEAN=mean(ext.streak_10gen.median, na.rm=TRUE), #mean of median length of streaks +10gen / replicate, na.rm=TRUE 

    # DATA WITH ANY OUTLIER STREAK LENGTHS REMOVED
    # count of replicates with extreme streaks present (NA-n, NAs present)
    gen3.nreps.streak.NOos.PRESENT=sum(!is.na(ext.streak_3gen.n.noouts)), #count n replicates with any streaks more than 3 gen (where not NA) 
    gen5.nreps.streak.NOos.PRESENT=sum(!is.na(ext.streak_5gen.n.noouts)), #count n replicates with any streaks more than 5 gen (where not NA) 
    gen10.nreps.streak.NOos.PRESENT=sum(!is.na(ext.streak_10gen.n.noouts)), #count n replicates with any streaks more than 10 gen (where not NA) 
    
    # mean of number of streaks / replicate, na.rm=TRUE
    gen3.NstreakXrep.NOos.MEAN=mean(ext.streak_3gen.n.noouts, na.rm=TRUE), #streaks more than 3 gen
    gen5.NstreakXrep.NOos.MEAN=mean(ext.streak_5gen.n.noouts, na.rm=TRUE), #streaks more than 5 gen
    gen10.NstreakXrep.NOos.MEAN=mean(ext.streak_10gen.n.noouts, na.rm=TRUE), #streaks more than 10 gen
    
    # mean of median streak length per replicate (NA-n, NAs present)  
    gen3.medianstreaklenXrep.NOos.MEAN=mean(ext.streak_3gen.median.noouts, na.rm=TRUE), #streaks more than 3 gen
    gen5.medianstreaklenXrep.NOos.MEAN=mean(ext.streak_5gen.median.noouts, na.rm=TRUE), #streaks more than 5 gen
    gen10.medianstreaklenXrep.NOos.MEAN=mean(ext.streak_10gen.median.noouts, na.rm=TRUE) #streaks more than 10 gen
      
)


```

## Check preference summary
Where "NaN" then it's a summary like mean or median (na.rm=TRUE) where there were all NAs and no other values
```{r}
nrow(compiled.preferences)
#str(compiled.preferences)
#unique(compiled.preferences$gen3.NstreakXrep.Wos.MEAN)
#subset(compiled.preferences, ta.first.stabilized.gen.fuzzy_median!=tb.first.stabilized.gen.fuzzy_median)
```


# Format PREFERENCE data for plotting heatmaps
```{r, selected.summaries, cache=FALSE, fig.height=12, fig.width=24}
#glimpse(compiled.preferences)
#names(compiled.preferences)
#str(compiled.preferences)


#melt preference/trait data x model, setting, groupsize, s, a, thresholds
summary_across.replicates.preferences.mlt <- reshape2::melt(compiled.preferences, 
                                                id.vars = c("model", "setting", "groupn", "s", "a", "loop.thresholds"))

#rename measure column
summary_across.replicates.preferences.mlt <- summary_across.replicates.preferences.mlt%>% rename(measure=variable)

#add MS key for easier plotting
summary_across.replicates.preferences.mlt$key.MS <- paste(
                                                      summary_across.replicates.preferences.mlt$model, 
                                                      summary_across.replicates.preferences.mlt$setting, 
                                                      sep="."
                                                      )

glimpse(summary_across.replicates.preferences.mlt)

# add letter to more easily sory panels for plotting
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="fd.x"] <- "a"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="fd.fa"] <- "b"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="fd.fs"] <- "c"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="fde.x"] <- "d"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="fde.xpt1"] <- "e"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="ncs.x"] <- "f"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="pfd.x"] <- "g"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="pnov.x"] <- "h"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="sc.x"] <- "i"



summary_across.replicates.preferences.mlt$model.plotlab <- paste0("(", 
                                                                  summary_across.replicates.preferences.mlt$model.plotord, ") ", 
                                                                  summary_across.replicates.preferences.mlt$key.MS
                                                                  )
```

## Melt and format TRAIT data
```{r}
#names(compiled.traits)
#melt preference/trait data x model, setting, groupsize, s, a, thresholds
summary_across.replicates.traits.mlt <- reshape2::melt(compiled.ta, 
                                                id.vars = c("model", "setting", "groupn", "s", "a"))

#rename measure column
summary_across.replicates.traits.mlt <- summary_across.replicates.traits.mlt%>% rename(measure=variable)

#add MS key for easier plotting
summary_across.replicates.traits.mlt$key.MS <- paste(
                                                      summary_across.replicates.traits.mlt$model, 
                                                      summary_across.replicates.traits.mlt$setting, 
                                                      sep="."
                                                      )

#glimpse(summary_across.replicates.traits.mlt)

# add letter to more easily sory panels for plotting
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="fd.x"] <- "a"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="fd.fa"] <- "b"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="fd.fs"] <- "c"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="fde.x"] <- "d"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="fde.xpt1"] <- "e"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="ncs.x"] <- "f"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="pfd.x"] <- "g"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="pnov.x"] <- "h"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="sc.x"] <- "i"



summary_across.replicates.traits.mlt$model.plotlab <- paste0("(", 
                                                                  summary_across.replicates.traits.mlt$model.plotord, ") ", 
                                                                  summary_across.replicates.traits.mlt$key.MS
                                                                  )

glimpse(summary_across.replicates.traits.mlt)
```


## Make key of all measures to run
```{r}
#KEY to TRAIT measures
list.measures.TRAIT.summ <- as.data.frame(as.character(unique(summary_across.replicates.traits.mlt$measure)))
names(list.measures.TRAIT.summ) <- "measure"
list.measures.TRAIT.summ$to.run <- "Y"
summ.mms.TRAIT.torun <- subset(list.measures.TRAIT.summ, to.run=="Y") 

#KEY to PREFERENCE measures

#add threshold to measure to make plotting easier
summary_across.replicates.preferences.mlt$measure.threshold <- paste(summary_across.replicates.preferences.mlt$measure, summary_across.replicates.preferences.mlt$thresholds, sep=".") #head(summary_across.replicates.preferences.mlt$measure.threshold)

list.measures.PREF.summ <- as.data.frame(as.character(unique(summary_across.replicates.preferences.mlt$measure.threshold)))
names(list.measures.PREF.summ) <- "measure"
list.measures.PREF.summ$to.run <- "Y"
summ.mms.PREF.torun <- subset(list.measures.PREF.summ, to.run=="Y")


```

# Loop to plot heatmaps for each measure

## set colors
```{r}
female.cols <- brewer.pal(6, "OrRd")
male.cols <- brewer.pal(6, "GnBu")

pal_female <- colorRampPalette(female.cols)
pal_male <- colorRampPalette(male.cols)
```
## TRAIT data: stability
```{r}
#set up progress bar
n.loops<-length(summ.mms.TRAIT.torun$measure) # Number of iterations
pb <- txtProgressBar(min = 0, max = n.loops, style = 3) # Initiate progress bar


#start pdf plot
pdf(paste("Plots_summarized data/ALLheatmaps_traitsRun02", "pdf", sep="."), width=16, height=9, onefile = TRUE)

#loop to plot
for(m.s in 1:length(summ.mms.TRAIT.torun$measure)){ #m.s=1
  loop.mm <- as.character(summ.mms.TRAIT.torun$measure[m.s])
  mm.data <- subset(summary_across.replicates.traits.mlt, measure==loop.mm) #glimpse(mm.data)
  mm.data$value <- as.numeric(mm.data$value)
  
  # Update the progress bar
   #setTxtProgressBar(pb, m.s)
   
     #use factor and levels to force arrangement of plot
  mm.data$s.sort = factor(mm.data$s, levels=c(1, 0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.1, 0))
  mm.data$grp.sort = factor(mm.data$groupn, levels=c("10", "30", "100"))

  #set color palette
  pal <- colorRampPalette(rev(brewer.pal(11, 'Spectral')), space='Lab')

  # highlight cells (see https://www.py4u.net/discuss/862186)
  mm.data$highlight <- FALSE
  mm.data$highlight[mm.data$s==0.8 & mm.data$a==2 | mm.data$s==0.8 & mm.data$a==7 | mm.data$s==0.2 & mm.data$a==2 | mm.data$s==0.2 & mm.data$a==7] <- TRUE
  #unique(mm.data$highlight)
  cust.highlight.cells <- subset(mm.data, highlight==TRUE)
  cust.highlight.cells <- subset(cust.highlight.cells, select=c(grp.sort,model.plotlab, a,s,highlight))
  cust.highlight.cells$a.ajusted <- cust.highlight.cells$a+1
  cust.highlight.cells$s.ajusted <- (cust.highlight.cells$s*10)+1

  
  # geom_rect(data=cust.highlight.cells, size=1, fill=NA, colour="black",
    #              aes(xmin=a - 0.5, xmax=a + 0.5, ymin=s - 0.5, ymax=s + 0.5))

  #Plot faceted heatmaps
    heat <- ggplot(data = mm.data, aes(x=as.factor(a), y=as.factor(s), fill=value)) + 
              geom_tile(colour="white",size=0.25) +
              #scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
              labs(title = paste0("Summary values:", " Measure=", loop.mm), y = "s", x = "a") +
              scale_fill_gradientn(colours = pal(100)) +
              geom_rect(data=cust.highlight.cells, size=1, fill=NA, colour="black",
                  aes(xmin=as.numeric(a.ajusted) - 0.5, xmax=as.numeric(a.ajusted) + 0.5, ymin=as.numeric(s.ajusted) - 0.5, ymax=as.numeric(s.ajusted) + 0.5)) +
              theme(panel.background = element_blank(),
                          axis.line = element_line(colour = "black"),
                          axis.text=element_text(size=8),
                          #strip.text.x = element_blank(),
                          panel.grid.major = element_blank(), 
                          panel.grid.minor = element_blank(),
                          panel.border = element_rect(colour = "black", fill=NA, size=0.5),
                          plot.title = element_text(size=10),
                          plot.subtitle = element_text(size=8))
      

    print(heat + facet_grid(rows=vars(grp.sort), cols=vars(model.plotlab))) # set to model.plotlab to plot models in columns
}   

#to set color limits:
#scale_fill_gradientn(colours = pal(100), limits = c(0,20))

#end plotting (outside loop so that all plots are in single file)
dev.off()
```


# Plot preference and trait heatmaps
```{r}

#subset just group size 30 data from whole
summary_across.replicates.preferences.mlt.G30.25x75 <- subset(summary_across.replicates.preferences.mlt, groupn=="30" & loop.thresholds=="25_75")
summary_across.replicates.traits.mlt.G30 <- subset(summary_across.replicates.traits.mlt, groupn=="30")

unique(summary_across.replicates.preferences.mlt.G30.25x75$measure)

unique.MS.keys <- unique(summary_across.replicates.preferences.mlt.G30.25x75$key.MS)


#start pdf plot
pdf(paste("Plots_summarized data/ALLheatmaps_preferences&traitsRun02s", "pdf", sep="."), width=12, height=4.5, onefile = TRUE)

#loop to plot
for(KEY in 1:length(unique.MS.keys)){ #KEY=4
  loop.ms.key <- as.character(unique.MS.keys[KEY])
  
  #extract trait data for this loop's model and setting combo
  loop.data_PREF <- subset(summary_across.replicates.preferences.mlt.G30.25x75, key.MS==loop.ms.key) 
  loop.data_PREF$value <- as.numeric(loop.data_PREF$value)
  
  #extract trait data for this loop's model and setting combo
  loop.data_TRAIT <- subset(summary_across.replicates.traits.mlt.G30, key.MS==loop.ms.key) 
  loop.data_TRAIT$value <- as.numeric(loop.data_TRAIT$value)
  
  #extract example time series
  loop.ts.exs <- subset(example.ts.melt, key.MS==loop.ms.key) #tail(loop.ts.exs, 50)
  loop.ts.exs_p <- subset(loop.ts.exs, param=="p")
  loop.ts.exs_ta <- subset(loop.ts.exs, param=="ta")
  loop.ts.exs_tb <- subset(loop.ts.exs, param=="tb")
  
  
  
  #extract generation at fixation for vertical line
  loop.ta.fixation <- subset(run02s_trait.summ.sa, key.MS==loop.ms.key)
  
  # format PREFERENCE data for colors and labels in plot
      pref.summ_color <- subset(loop.data_PREF, measure=="extcircuit.nreps.PRESENT")
      pref.summ_color <- subset(pref.summ_color, select=c(key.MS, model.plotlab, groupn, s, a, loop.thresholds, measure, value))
      names(pref.summ_color)[names(pref.summ_color) == 'value'] <- 'extcircuit.nreps.PRESENT'
      names(pref.summ_color)[names(pref.summ_color) == 'measure'] <- 'color'
      
      pref.summ_label <- subset(loop.data_PREF, measure=="nshiftsXrep.MEDIAN")
      pref.summ_label <- subset(pref.summ_label, select=c(key.MS, model.plotlab, groupn, s, a, loop.thresholds, measure, value))
      names(pref.summ_label)[names(pref.summ_label) == 'value'] <- 'nshiftsXrep.MEDIAN'
      names(pref.summ_label)[names(pref.summ_label) == 'measure'] <- 'label'
      
  # format TRIAT data for colors and labels in plot
      trait.summ_color <- subset(loop.data_TRAIT, measure=="ta.nruns.variation.maintained_fuzzy")
      trait.summ_color <- subset(trait.summ_color, select=c(key.MS, model.plotlab, groupn, s, a, measure, value))
      names(trait.summ_color)[names(trait.summ_color) == 'value'] <- 'ta.nruns.variation.maintained_fuzzy'
      names(trait.summ_color)[names(trait.summ_color) == 'measure'] <- 'color'
      trait.summ_label <- subset(loop.data_TRAIT, measure=="ta.first.stabilized.gen.fuzzy_median")
      trait.summ_label <- subset(trait.summ_label, select=c(key.MS, model.plotlab, groupn, s, a, measure, value))
      names(trait.summ_label)[names(trait.summ_label) == 'value'] <- 'ta.first.stabilized.gen.fuzzy_median'
      names(trait.summ_label)[names(trait.summ_label) == 'measure'] <- 'label'  


#merge color and label data
pref.summ <- merge(pref.summ_color, 
                         pref.summ_label, 
                         by=c("key.MS" ,"model.plotlab", "groupn", "s", "a", "loop.thresholds"))

trait.summ <- merge(trait.summ_color, 
                         trait.summ_label, 
                         by=c("key.MS" ,"model.plotlab", "groupn", "s", "a"))

# highlight cells (see https://www.py4u.net/discuss/862186)
  pref.summ$highlight <- FALSE
  pref.summ$highlight[pref.summ$s==0.4 & pref.summ$a==2 | pref.summ$s==0.4 & pref.summ$a==7 | pref.summ$s==0.1 & pref.summ$a==2 | pref.summ$s==0.1 & pref.summ$a==7] <- TRUE
  #unique(mm.data$highlight)
  cust.highlight.cells <- subset(pref.summ, highlight==TRUE)
  cust.highlight.cells <- subset(cust.highlight.cells, select=c(groupn,model.plotlab, a,s,highlight))
  cust.highlight.cells$a.ajusted <- cust.highlight.cells$a+1
  cust.highlight.cells$s.ajusted <- (cust.highlight.cells$s*10)+1
  

  #use factor and levels to force arrangement of plot
  pref.summ$s.sort = factor(pref.summ$s, levels=c(0.4, 0.1))
  trait.summ$s.sort = factor(trait.summ$s, levels=c(0.4, 0.1))

    #sort s as factor for the two examples of s
  loop.ts.exs_tb$s = factor(loop.ts.exs_tb$s, levels=c(0.4, 0.1))
  loop.ts.exs_ta$s = factor(loop.ts.exs_ta$s, levels=c(0.4, 0.1))
  loop.ts.exs_p$s = factor(loop.ts.exs_p$s, levels=c(0.4, 0.1))
  loop.ta.fixation$s = factor(loop.ta.fixation$s, levels=c(0.4, 0.1))
  #set page title
  pg.title <- paste0("Model & setting: ", loop.ms.key," (group size=30)")

# EXAMPLE TIME SERIES
  
  ts.legend.cols <- c("Females preferring Trait B"="red2", "Frequency of male trait TA2"="royalblue3","Frequency of male trait TB2"="skyblue")

  ex.ts.plots <- ggplot() + 
                    ylim(0,1) +
                    xlim(1,100) +
                    geom_rect(data=loop.ta.fixation, 
                              mapping=aes(xmin=1, xmax=100, ymin=0.25, ymax=0.75), 
                              fill="lightgrey", color=NA, alpha=0.5) +
                    geom_hline(yintercept=0.5, color="darkgrey") +
                    geom_line(data=loop.ts.exs_tb, 
                              aes(x=gen, y=value, color="Frequency of male trait TB2"), size=0.5) +
                    geom_line(data=loop.ts.exs_ta, 
                              aes(x=gen, y=value, color="Frequency of male trait TA2"), size=0.5) + 
                    geom_line(data=loop.ts.exs_p, 
                              aes(x=gen, y=value, color="Females preferring Trait B"), size=0.65) +
                    scale_colour_manual(name="Parameters: ",values=ts.legend.cols)+
                    labs(title=pg.title, 
                         y="Prevalence in population", x="Generation",
                         subtitle="(a) Examples of dynamics of female preference and male traits") +
                    theme(panel.background = element_blank(),
                          axis.line = element_line(colour = "black"),
                          axis.text=element_text(size=8),
                          panel.grid.major = element_blank(), 
                          panel.grid.minor = element_blank(),
                          panel.border = element_rect(colour = "black", fill=NA, size=1.25),
                          plot.title = element_text(size=10),
                          plot.subtitle = element_text(size=9),
                          legend.position="bottom", legend.direction="vertical",
                          legend.text=element_text(size=8),
                          legend.key = element_rect(fill = "white"),
                          legend.title = element_blank()) +
                   facet_grid(s~a, #factor(s, levels=c(0.8,0.2))
                              labeller = label_both) +
                   geom_vline(data=loop.ta.fixation, 
                               aes(xintercept=first.stable.gen.fuzzy), 
                               color="royalblue3", linetype="dashed") 
  #print(ex.ts.plots)
                  
  
  #PREFERENCE HEATMAP (threshold=25x75, group size=30)
    heat.pref <- ggplot(data = pref.summ, aes(x=as.factor(a), y=as.factor(s), fill=extcircuit.nreps.PRESENT)) + 
              geom_tile(colour="white",size=0.25) +
              # cell labels
              geom_text(aes(label = round(nshiftsXrep.MEDIAN, 1), 
                                    color = extcircuit.nreps.PRESENT>9), size=2.5) +
              scale_color_manual(guide = FALSE, values = c("black", "white")) +
              #scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
              labs(title = pg.title, 
                   subtitle="(b) Extreme switches in female preferences",
                   y = "Strength of viability selection (s)", 
                   x = expression(paste("Strength of sexual selection (", alpha, ")"))) +
              scale_fill_gradientn(colours = c("white", pal_female(100)), 
                                   limits = c(0,20), name = "Replicates with extreme circuit present",
                                   guide = guide_colorbar(frame.colour = "black", ticks.colour = "black")) +
              geom_rect(data=cust.highlight.cells, size=1, fill=NA, colour="black",
                  aes(xmin=as.numeric(a.ajusted) - 0.5, xmax=as.numeric(a.ajusted) + 0.5, ymin=as.numeric(s.ajusted) - 0.5, ymax=as.numeric(s.ajusted) + 0.5)) +
       theme(panel.background = element_blank(),
                          axis.line = element_line(colour = "black"),
                          axis.text=element_text(size=8),
                          #strip.text.x = element_blank(),
                          panel.grid.major = element_blank(), 
                          panel.grid.minor = element_blank(),
                          panel.border = element_rect(colour = "black", fill=NA, size=0.5),
                          plot.title = element_text(size=10),
                          plot.subtitle = element_text(size=9),
                          legend.position="bottom", legend.direction="horizontal",
                          legend.text=element_text(size=8),
                          legend.title=element_text(size=8),
                          aspect.ratio = 1)
      
    #print(heat.pref) 
    
    #TRAIT HEATMAP (group size=30)
    heat.trait <- ggplot(data = trait.summ, aes(x=as.factor(a), y=as.factor(s), fill=ta.nruns.variation.maintained_fuzzy)) + 
              geom_tile(colour="white",size=0.25) +
              # cell labels
              geom_text(aes(label = round(ta.first.stabilized.gen.fuzzy_median, 1), 
                                    color = ta.nruns.variation.maintained_fuzzy>15), size=2.5) +
              scale_color_manual(guide = FALSE, values = c("black", "white")) +
              #scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
              labs(title = pg.title, 
                   subtitle="(c) Maintenance of genetic variation in male trait",
                   y = "Strength of viability selection (s)", 
                   x = expression(paste("Strength of sexual selection (", alpha, ")")))  +
              scale_fill_gradientn(colours = c("white", pal_male(100)), 
                                   limits = c(0,20), name = "Replicates retaining trait variability", 
                                   guide = guide_colorbar(frame.colour = "black", ticks.colour = "black")) +
              geom_rect(data=cust.highlight.cells, size=1, fill=NA, colour="black",
                  aes(xmin=as.numeric(a.ajusted) - 0.5, xmax=as.numeric(a.ajusted) + 0.5, ymin=as.numeric(s.ajusted) - 0.5, ymax=as.numeric(s.ajusted) + 0.5)) +
       theme(panel.background = element_blank(),
                          axis.line = element_line(colour = "black"),
                          axis.text=element_text(size=8),
                          #strip.text.x = element_blank(),
                          panel.grid.major = element_blank(), 
                          panel.grid.minor = element_blank(),
                          panel.border = element_rect(colour = "black", fill=NA, size=0.5),
                          plot.title = element_text(size=10),
                          plot.subtitle = element_text(size=9),
                          legend.position="bottom", legend.direction="horizontal",
                          legend.text=element_text(size=8),
                          legend.title=element_text(size=8),
                          aspect.ratio = 1)
      
    #print(heat.trait) 
    
#pdf(paste("Fig_compiled_fd.fa", "pdf", sep="."), width=16, height=7, onefile = TRUE)
gridExtra::grid.arrange(ex.ts.plots, heat.pref, heat.trait, nrow = 1, ncol=3)  

}   
    
    
    
#end plotting (outside loop so that all plots are in single file)
dev.off()


```
