---
title: "Step04_summarize tseries preferences & traits for heatmaps"
author: "Liz Hobson"
date: "2022-10-02"
output: html_document
---

This script is Step 4a of the analysis/plotting scripts. 

In this script we summarize patterns in preferences and trait variation for each s/a combination, for each group size, and for each model type. These summaries are used to plot the heat maps seen in the main text in the figures Fig. 1 (c/d), Fig. 3 (aii/aiii and bii/biii), and Fig. 4 (b/c). These summaries are also used to plot the heat maps seen in the supplemental figures Fig. S1 (aii/aiii, bii/biii, and cii/ciii). This script also pulls out four exemplar raw timeseries runs and plots them. These appear in the main text as panels in Fig. 1(b), Fig. 3 (ai/bi), and Fig. 4(a, but Step4b for one exception). These time series exemplars also appear in the supplemental in Figure S1(ai/bi/ci).


# Import data & load packages
```{r}
library(dplyr)
library(tidyverse)
library(reshape2)
library(gplots)
library(RColorBrewer)
library(extraoperators)
library(stringr)          

cv <- function(x) 100*( sd(x)/mean(x))

# set data path
compiled.data.path <- "Cached_data/"

file.list <- as.data.frame(list.files(compiled.data.path, recursive=TRUE))
colnames(file.list) <- "file.name"

# extract just file names with fixation
file.list_trait.fixation <- as.data.frame(file.list[stringr::str_detect(file.list$file.name, "fixation"), ])  # Extract matching rows with str_detect
colnames(file.list_trait.fixation) <- "file.name"
#compile trait data
data_traits <- do.call(rbind, lapply(paste0(compiled.data.path,file.list_trait.fixation$file.name), function(x) read.csv(x, stringsAsFactors = FALSE)))


# extract just file names with preference
file.list_preferences <- as.data.frame(file.list[stringr::str_detect(file.list$file.name, "preference"), ]) # Extract matching rows with str_detect
colnames(file.list_preferences) <- "file.name"
#compile preference data
data_preferences <- do.call(rbind, lapply(paste0(compiled.data.path,file.list_preferences$file.name), function(x) read.csv(x, stringsAsFactors = FALSE)))

```



# Data wrangling
```{r}
# sort
data_traits <- data_traits %>% arrange(model, setting, groupn,s,a, run) #names(data_traits)
data_preferences <- data_preferences %>% arrange(model, setting, groupn,s,a, run, loop.thresholds) 
```

# Summarize trait data for Trait A
Find the median generation at which fixation occurred for all runs where fixation did occur (exclude any runs where variation was maintained)
```{r}
unique(data_traits$first.stable.gen.fuzzy)

compiled.ta <- data_traits %>% 
                  dplyr::group_by(model, setting, groupn, s, a) %>% 
                  dplyr::summarize(
                      # for strict stabilization definition
                        ta.stabilized_Nruns=sum(first.stable.gen<100), 
                        ta.first.stabilized.gen_median=median(first.stable.gen[first.stable.gen<100]),
                      # for fuzzy stabilization definition
                        ta.runs.stabilized.fuzzy_N=sum(first.stable.gen.fuzzy<100), #how many runs n.gens.stable>0 (NA if it didn't stabilize)
                        ta.first.stabilized.gen.fuzzy_median=median(first.stable.gen.fuzzy[first.stable.gen.fuzzy<100]),
                      ta.nruns.variation.maintained_fuzzy=20-ta.runs.stabilized.fuzzy_N
                )

#glimpse(compiled.ta)
unique(compiled.ta$ta.nruns.variation.maintained_fuzzy)

```

#Import raw timeseries data & rbind together
```{r}
# set data path
raw.data.path <- "Rawdata_compiled/"

# import data
runs.fd.fa <- read.csv(paste0(raw.data.path, "compiledruns.fd.fa.csv"))  
runs.fd.fs <- read.csv(paste0(raw.data.path, "compiledruns.fd.fs.csv")) 
runs.fd.x <- read.csv(paste0(raw.data.path, "compiledruns.fd.x.csv")) 
runs.fde.x <- read.csv(paste0(raw.data.path, "compiledruns.fde.x.csv")) 
runs.fde.xpt1 <- read.csv(paste0(raw.data.path, "compiledruns.fde.xpt1.csv"))  
runs.ncs.x <- read.csv(paste0(raw.data.path, "compiledruns.ncs.x.csv")) 
runs.pfd.x <- read.csv(paste0(raw.data.path, "compiledruns.pfd.x.csv")) 
runs.pnov.x <- read.csv(paste0(raw.data.path, "compiledruns.pnov.x.csv"))  
runs.sc.x <- read.csv(paste0(raw.data.path, "compiledruns.sc.x.csv")) 

# compile data
all.runs <- rbind.data.frame(runs.fd.fa, 
                             runs.fd.fs,
                             runs.fd.x, 
                             runs.fde.x,
                             runs.fde.xpt1,
                             runs.ncs.x,
                             runs.pfd.x,
                             runs.pnov.x,
                             runs.sc.x
                             ) 
# check compiled data
glimpse(all.runs)

# runs have rowIDs, delete that first column
all.runs <- select(all.runs, -1) #glimpse(all.runs)
glimpse(all.runs)
```

## Extract metadata from raw timeseries files
```{r}
#separate parts of file names
# for col 1 which has full names, split it out then split the parts
all.runs.split.names <- all.runs[,1] #head(all.runs[,1]) #head(all.runs.split.names)

all.runs.split.names <- strsplit(as.character(all.runs.split.names), ".", fixed = TRUE) #head(all.runs.split.names)

#unlist data from one rowxcolumn cell
all.runs.split.names <- as.data.frame(do.call(rbind, all.runs.split.names)) #head(all.runs.split.names) #str(all.runs.split.names)

# add informative column names
colnames(all.runs.split.names) <- c("param", "model", "setting", "group.size", "run") #head(all.runs.split.names)

all.runs.split.names$runID <- sub('.', '', all.runs.split.names$run)
all.runs.split.names$group.size <- as.numeric(sub('.', '', all.runs.split.names$group.size)) 

all.runs.split.names$runID <- as.numeric(all.runs.split.names$runID) #str(all.runs.split.names) unique(all.runs.split.names$setting)


#bind back to all.runs
all.runsWmeta <- cbind.data.frame(all.runs.split.names, all.runs) #str(all.runsWmeta) #str(all.runs)

#add individual run ID name
all.runsWmeta$fileWsa <- paste0(all.runsWmeta$file, "_s", all.runsWmeta$s, "_a", all.runsWmeta$a)



all.runsWmeta <- all.runsWmeta %>% select(-run, -file, -fileWsa) #str(all.runsWmeta) length(all.runsWmeta$runID)
unique(all.runsWmeta$setting)
unique(all.runsWmeta$param)
names(all.runsWmeta)
unique(all.runsWmeta$group.size)

glimpse(all.runsWmeta)
head(all.runsWmeta)
```

### CHECK OUTPUT OF RUNS
Should be 110 subparts per model X params X N X run number (and 2200 per model X params X N)
```{r, echo=TRUE}
run.meta.summary <- all.runsWmeta %>% dplyr::group_by(param, model, setting, group.size) %>% tally() 

#run.meta.summary
unique(run.meta.summary$n) #all of these should have 2200 runs each

```


# Import shortcut data
```{r}
all.runsWmeta.params_ta.tb.p <- subset(all.runsWmeta, param=="ta" | param=="tb" | param=="p")
example.ts <- subset(all.runsWmeta.params_ta.tb.p, runID==1 & group.size==30)
example.ts <- subset(example.ts, s==0.1 | s==0.4)
example.ts <- subset(example.ts, a==2 | a==7)
names(example.ts)
example.ts.melt <- reshape2::melt(example.ts, id.vars = c("model", "setting", "group.size", "param", "runID", "s", "a"))
head(example.ts.melt)
#rename measure column
example.ts.melt$gen <- as.numeric(sub('.', '', example.ts.melt$variable))
example.ts.melt <- example.ts.melt %>% arrange(model, setting, group.size, param, s, a, runID, gen)
#add MS key for easier plotting
example.ts.melt$key.MS <- paste(example.ts.melt$model, 
                                example.ts.melt$setting, 
                                sep="."
                                )


# Retain just the summary data for run 1
run01s_trait.summ <- subset(data_traits, run=="1" & groupn==30)
run01s_trait.summ.sa <- subset(run01s_trait.summ, s==0.1 | s==0.4)
run01s_trait.summ.sa <- subset(run01s_trait.summ.sa, a==2 | a==7)
#add MS key for easier plotting
run01s_trait.summ.sa$key.MS <- paste(run01s_trait.summ.sa$model, 
                                run01s_trait.summ.sa$setting, 
                                sep="."
                                )
names(run01s_trait.summ.sa) #use first.stable.gen.fuzzy to plot

```

# Summarize preference data 
```{r}

#str(streak_p)
compiled.preferences <- data_preferences %>% dplyr::group_by(model, setting, groupn, s, a, loop.thresholds) %>% 
  dplyr::summarize(
    # median number of values in lower/upper thresholds
    n_lower.median=median(n_lower, na.rm=TRUE), 
    n_upper.median=median(n_upper, na.rm=TRUE), 
    
    # number extreme shifts (1=went from one extreme to other once, no NA, >=2 at least one cycle)
    extcircuit.nreps.PRESENT=sum(n.extreme.shifts>=2), #count n replicates with shifts >=2 (at least one "cycle")
    nshiftsXrep.MEDIAN=median(n.extreme.shifts), #median across all replicates, INCLUDING 0
    nshiftsXrep.MEDIAN.no0=median(n.extreme.shifts[n.extreme.shifts>=2]), #median across all replicates, EXCLUDING 0
    
    # # count number runs with outliers (0-n, no NAs)
    # outliers.n.PRESENT=sum(outliers.n>0), #n replicates with any outliers (count where >0)
    # outliers.n.MEDIAN=median(outliers.n), #median number of outliers per replicate, including 0
    
    # ALL DATA, INCLUDING OUTINER STREAK LENGTHS
    # count of replicates with extreme streaks present (NA-n, NAs present)
    gen3.nreps.streak.Wos.PRESENT=sum(!is.na(ext.streak_3gen.n)), #count n replicates with any streaks more than 3 gen (where not NA) 
    gen5.nreps.streak.Wos.PRESENT=sum(!is.na(ext.streak_5gen.n)), #count n replicates with any streaks more than 5 gen (where not NA) 
    gen10.nreps.streak.Wos.PRESENT=sum(!is.na(ext.streak_10gen.n)), #count n replicates with any streaks more than 10 gen (where not NA) 
    
    # mean number of streaks per replicate (NA-n, NAs present)
    gen3.NstreakXrep.Wos.MEAN=mean(ext.streak_3gen.n, na.rm=TRUE),#mean of number of streaks / replicate, na.rm=TRUE (streaks more than 3 gen)
    gen5.NstreakXrep.Wos.MEAN=mean(ext.streak_5gen.n, na.rm=TRUE), #mean of number of streaks / replicate, na.rm=TRUE (streaks more than 5 gen)
    gen10.NstreakXrep.Wos.MEAN=mean(ext.streak_10gen.n, na.rm=TRUE), #mean of number of streaks / replicate, na.rm=TRUE (streaks more than 10 gen)
    
    # mean of median streak length per replicate (NA-n, NAs present)  
    gen3.medianstreaklenXrep.Wos.MEAN=mean(ext.streak_3gen.median, na.rm=TRUE), #mean of median length of streaks +3gen / replicate, na.rm=TRUE 
    gen5.medianstreaklenXrep.Wos.MEAN=mean(ext.streak_5gen.median, na.rm=TRUE), #mean of median length of streaks +5gen / replicate, na.rm=TRUE 
    gen10.medianstreaklenXrep.Wos.MEAN=mean(ext.streak_10gen.median, na.rm=TRUE), #mean of median length of streaks +10gen / replicate, na.rm=TRUE 

    # DATA WITH ANY OUTLIER STREAK LENGTHS REMOVED
    # count of replicates with extreme streaks present (NA-n, NAs present)
    gen3.nreps.streak.NOos.PRESENT=sum(!is.na(ext.streak_3gen.n.noouts)), #count n replicates with any streaks more than 3 gen (where not NA) 
    gen5.nreps.streak.NOos.PRESENT=sum(!is.na(ext.streak_5gen.n.noouts)), #count n replicates with any streaks more than 5 gen (where not NA) 
    gen10.nreps.streak.NOos.PRESENT=sum(!is.na(ext.streak_10gen.n.noouts)), #count n replicates with any streaks more than 10 gen (where not NA) 
    
    # mean of number of streaks / replicate, na.rm=TRUE
    gen3.NstreakXrep.NOos.MEAN=mean(ext.streak_3gen.n.noouts, na.rm=TRUE), #streaks more than 3 gen
    gen5.NstreakXrep.NOos.MEAN=mean(ext.streak_5gen.n.noouts, na.rm=TRUE), #streaks more than 5 gen
    gen10.NstreakXrep.NOos.MEAN=mean(ext.streak_10gen.n.noouts, na.rm=TRUE), #streaks more than 10 gen
    
    # mean of median streak length per replicate (NA-n, NAs present)  
    gen3.medianstreaklenXrep.NOos.MEAN=mean(ext.streak_3gen.median.noouts, na.rm=TRUE), #streaks more than 3 gen
    gen5.medianstreaklenXrep.NOos.MEAN=mean(ext.streak_5gen.median.noouts, na.rm=TRUE), #streaks more than 5 gen
    gen10.medianstreaklenXrep.NOos.MEAN=mean(ext.streak_10gen.median.noouts, na.rm=TRUE) #streaks more than 10 gen
      
)


```

## Check preference summary
Where "NaN" then it's a summary like mean or median (na.rm=TRUE) where there were all NAs and no other values
```{r}
nrow(compiled.preferences)
#str(compiled.preferences)
#unique(compiled.preferences$gen3.NstreakXrep.Wos.MEAN)
#subset(compiled.preferences, ta.first.stabilized.gen.fuzzy_median!=tb.first.stabilized.gen.fuzzy_median)
```


# Format PREFERENCE data for plotting heatmaps
```{r, selected.summaries, cache=FALSE, fig.height=12, fig.width=24}
#glimpse(compiled.preferences)
#names(compiled.preferences)
#str(compiled.preferences)


#melt preference/trait data x model, setting, groupsize, s, a, thresholds
summary_across.replicates.preferences.mlt <- reshape2::melt(compiled.preferences, 
                                                id.vars = c("model", "setting", "groupn", "s", "a", "loop.thresholds"))

#rename measure column
summary_across.replicates.preferences.mlt <- summary_across.replicates.preferences.mlt%>% rename(measure=variable)

#add MS key for easier plotting
summary_across.replicates.preferences.mlt$key.MS <- paste(
                                                      summary_across.replicates.preferences.mlt$model, 
                                                      summary_across.replicates.preferences.mlt$setting, 
                                                      sep="."
                                                      )

# add letter to more easily sort panels for plotting
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="fd.x"] <- "a"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="fd.fa"] <- "b"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="fd.fs"] <- "c"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="fde.x"] <- "d"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="fde.xpt1"] <- "e"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="ncs.x"] <- "f"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="pfd.x"] <- "g"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="pnov.x"] <- "h"
summary_across.replicates.preferences.mlt$model.plotord[summary_across.replicates.preferences.mlt$key.MS=="sc.x"] <- "i"



summary_across.replicates.preferences.mlt$model.plotlab <- paste0("(", 
                                                                  summary_across.replicates.preferences.mlt$model.plotord, ") ", 
                                                                  summary_across.replicates.preferences.mlt$key.MS
                                                                  )

glimpse(summary_across.replicates.preferences.mlt)

```

## Melt and format TRAIT data
```{r}
#names(compiled.traits)
#melt preference/trait data x model, setting, groupsize, s, a, thresholds
summary_across.replicates.traits.mlt <- reshape2::melt(compiled.ta, 
                                                id.vars = c("model", "setting", "groupn", "s", "a"))

#rename measure column
summary_across.replicates.traits.mlt <- summary_across.replicates.traits.mlt%>% rename(measure=variable)

#add MS key for easier plotting
summary_across.replicates.traits.mlt$key.MS <- paste(
                                                      summary_across.replicates.traits.mlt$model, 
                                                      summary_across.replicates.traits.mlt$setting, 
                                                      sep="."
                                                      )


# add letter to more easily sort panels for plotting
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="fd.x"] <- "a"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="fd.fa"] <- "b"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="fd.fs"] <- "c"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="fde.x"] <- "d"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="fde.xpt1"] <- "e"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="ncs.x"] <- "f"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="pfd.x"] <- "g"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="pnov.x"] <- "h"
summary_across.replicates.traits.mlt$model.plotord[summary_across.replicates.traits.mlt$key.MS=="sc.x"] <- "i"



summary_across.replicates.traits.mlt$model.plotlab <- paste0("(", 
                                                                  summary_across.replicates.traits.mlt$model.plotord, ") ", 
                                                                  summary_across.replicates.traits.mlt$key.MS
                                                                  )
glimpse(summary_across.replicates.traits.mlt)
```


## Make key of all measures to run
```{r}
#KEY to TRAIT measures
list.measures.TRAIT.summ <- as.data.frame(as.character(unique(summary_across.replicates.traits.mlt$measure)))
names(list.measures.TRAIT.summ) <- "measure"
list.measures.TRAIT.summ$to.run <- "Y"
summ.mms.TRAIT.torun <- subset(list.measures.TRAIT.summ, to.run=="Y") 

#KEY to PREFERENCE measures

#add threshold to measure to make plotting easier
summary_across.replicates.preferences.mlt$measure.threshold <- paste(summary_across.replicates.preferences.mlt$measure, summary_across.replicates.preferences.mlt$thresholds, sep=".") #head(summary_across.replicates.preferences.mlt$measure.threshold)

list.measures.PREF.summ <- as.data.frame(as.character(unique(summary_across.replicates.preferences.mlt$measure.threshold)))
names(list.measures.PREF.summ) <- "measure"
list.measures.PREF.summ$to.run <- "Y"
summ.mms.PREF.torun <- subset(list.measures.PREF.summ, to.run=="Y")


```



# Plot preference and trait heatmaps with exemplar time series
```{r}

#subset just group size 30 data from whole
summary_across.replicates.preferences.mlt.G30.25x75 <- subset(summary_across.replicates.preferences.mlt, groupn=="30" & loop.thresholds=="25_75")
summary_across.replicates.traits.mlt.G30 <- subset(summary_across.replicates.traits.mlt, groupn=="30")

unique(summary_across.replicates.preferences.mlt.G30.25x75$measure)

unique.MS.keys <- unique(summary_across.replicates.preferences.mlt.G30.25x75$key.MS)


#start pdf plot
#pdf(paste("Plots_summarized data/ALLheatmaps_preferences&traits", "pdf", sep="."), width=12, height=4.5, onefile = TRUE)

#loop to plot
for(KEY in 1:length(unique.MS.keys)){ #KEY=4
  loop.ms.key <- as.character(unique.MS.keys[KEY])
  
  #extract trait data for this loop's model and setting combo
  loop.data_PREF <- subset(summary_across.replicates.preferences.mlt.G30.25x75, key.MS==loop.ms.key) 
  loop.data_PREF$value <- as.numeric(loop.data_PREF$value)
  
  #extract trait data for this loop's model and setting combo
  loop.data_TRAIT <- subset(summary_across.replicates.traits.mlt.G30, key.MS==loop.ms.key) 
  loop.data_TRAIT$value <- as.numeric(loop.data_TRAIT$value)
  
  #extract example time series
  loop.ts.exs <- subset(example.ts.melt, key.MS==loop.ms.key) #tail(loop.ts.exs, 50)
  loop.ts.exs_p <- subset(loop.ts.exs, param=="p")
  loop.ts.exs_ta <- subset(loop.ts.exs, param=="ta")
  loop.ts.exs_tb <- subset(loop.ts.exs, param=="tb")
  
  
  
  #extract generation at fixation for vertical line
  loop.ta.fixation <- subset(run01s_trait.summ.sa, key.MS==loop.ms.key)
  
  # format PREFERENCE data for colors and labels in plot
      pref.summ_color <- subset(loop.data_PREF, measure=="extcircuit.nreps.PRESENT")
      pref.summ_color <- subset(pref.summ_color, select=c(key.MS, model.plotlab, groupn, s, a, loop.thresholds, measure, value))
      names(pref.summ_color)[names(pref.summ_color) == 'value'] <- 'extcircuit.nreps.PRESENT'
      names(pref.summ_color)[names(pref.summ_color) == 'measure'] <- 'color'
      
      pref.summ_label <- subset(loop.data_PREF, measure=="nshiftsXrep.MEDIAN")
      pref.summ_label <- subset(pref.summ_label, select=c(key.MS, model.plotlab, groupn, s, a, loop.thresholds, measure, value))
      names(pref.summ_label)[names(pref.summ_label) == 'value'] <- 'nshiftsXrep.MEDIAN'
      names(pref.summ_label)[names(pref.summ_label) == 'measure'] <- 'label'
      
  # format TRIAT data for colors and labels in plot
      trait.summ_color <- subset(loop.data_TRAIT, measure=="ta.nruns.variation.maintained_fuzzy")
      trait.summ_color <- subset(trait.summ_color, select=c(key.MS, model.plotlab, groupn, s, a, measure, value))
      names(trait.summ_color)[names(trait.summ_color) == 'value'] <- 'ta.nruns.variation.maintained_fuzzy'
      names(trait.summ_color)[names(trait.summ_color) == 'measure'] <- 'color'
      trait.summ_label <- subset(loop.data_TRAIT, measure=="ta.first.stabilized.gen.fuzzy_median")
      trait.summ_label <- subset(trait.summ_label, select=c(key.MS, model.plotlab, groupn, s, a, measure, value))
      names(trait.summ_label)[names(trait.summ_label) == 'value'] <- 'ta.first.stabilized.gen.fuzzy_median'
      names(trait.summ_label)[names(trait.summ_label) == 'measure'] <- 'label'  


#merge color and label data
pref.summ <- merge(pref.summ_color, 
                         pref.summ_label, 
                         by=c("key.MS" ,"model.plotlab", "groupn", "s", "a", "loop.thresholds"))

trait.summ <- merge(trait.summ_color, 
                         trait.summ_label, 
                         by=c("key.MS" ,"model.plotlab", "groupn", "s", "a"))

# highlight cells (see https://www.py4u.net/discuss/862186)
  pref.summ$highlight <- FALSE
  pref.summ$highlight[pref.summ$s==0.4 & pref.summ$a==2 | pref.summ$s==0.4 & pref.summ$a==7 | pref.summ$s==0.1 & pref.summ$a==2 | pref.summ$s==0.1 & pref.summ$a==7] <- TRUE
  #unique(mm.data$highlight)
  cust.highlight.cells <- subset(pref.summ, highlight==TRUE)
  cust.highlight.cells <- subset(cust.highlight.cells, select=c(groupn,model.plotlab, a,s,highlight))
  cust.highlight.cells$a.ajusted <- cust.highlight.cells$a+1
  cust.highlight.cells$s.ajusted[cust.highlight.cells$s==0.1] <- 3
  cust.highlight.cells$s.ajusted[cust.highlight.cells$s==0.4] <- 9
 

  #use factor and levels to force arrangement of plot
  pref.summ$s.sort = factor(pref.summ$s, levels=c(0.4, 0.1))
  trait.summ$s.sort = factor(trait.summ$s, levels=c(0.4, 0.1))

    #sort s as factor for the two examples of s
  loop.ts.exs_tb$s = factor(loop.ts.exs_tb$s, levels=c(0.4, 0.1))
  loop.ts.exs_ta$s = factor(loop.ts.exs_ta$s, levels=c(0.4, 0.1))
  loop.ts.exs_p$s = factor(loop.ts.exs_p$s, levels=c(0.4, 0.1))
  loop.ta.fixation$s = factor(loop.ta.fixation$s, levels=c(0.4, 0.1))
  #set page title
  pg.title <- paste0("Model & setting: ", loop.ms.key," (group size=30)")

# EXAMPLE TIME SERIES
  
  ts.legend.cols <- c("Females preferring Trait B"="red2", "Frequency of male trait TA2"="royalblue3","Frequency of male trait TB2"="skyblue")

  ex.ts.plots <- ggplot() + 
                    ylim(0,1) +
                    xlim(1,100) +
                    geom_rect(data=loop.ta.fixation, 
                              mapping=aes(xmin=1, xmax=100, ymin=0.25, ymax=0.75), 
                              fill="lightgrey", color=NA, alpha=0.5) +
                    geom_hline(yintercept=0.5, color="darkgrey") +
                    geom_line(data=loop.ts.exs_tb, 
                              aes(x=gen, y=value, color="Frequency of male trait TB2"), size=0.5) +
                    geom_line(data=loop.ts.exs_ta, 
                              aes(x=gen, y=value, color="Frequency of male trait TA2"), size=0.5) + 
                    geom_line(data=loop.ts.exs_p, 
                              aes(x=gen, y=value, color="Females preferring Trait B"), size=0.65) +
                    scale_colour_manual(name="Parameters: ",values=ts.legend.cols)+
                    labs(title=pg.title, 
                         y="Prevalence in population", x="Generation",
                         subtitle="(a) Examples of dynamics of female preference and male traits") +
                    theme(panel.background = element_blank(),
                          axis.line = element_line(colour = "black"),
                          axis.text=element_text(size=8),
                          panel.grid.major = element_blank(), 
                          panel.grid.minor = element_blank(),
                          panel.border = element_rect(colour = "black", fill=NA, size=1.25),
                          plot.title = element_text(size=10),
                          plot.subtitle = element_text(size=9),
                          legend.position="bottom", legend.direction="vertical",
                          legend.text=element_text(size=8),
                          legend.key = element_rect(fill = "white"),
                          legend.title = element_blank()) +
                   facet_grid(s~a, #factor(s, levels=c(0.8,0.2))
                              labeller = label_both) +
                   geom_vline(data=loop.ta.fixation, 
                               aes(xintercept=first.stable.gen.fuzzy), 
                               color="royalblue3", linetype="dashed") 
  #print(ex.ts.plots)
                  
  
  #PREFERENCE HEATMAP (threshold=25x75, group size=30)
    heat.pref <- ggplot(data = pref.summ, aes(x=as.factor(a), y=as.factor(s), fill=extcircuit.nreps.PRESENT)) + 
              geom_tile(colour="white",size=0.25) +
              # cell labels
              geom_text(aes(label = round(nshiftsXrep.MEDIAN, 1), 
                                    color = extcircuit.nreps.PRESENT>9), size=2.5) +
              scale_color_manual(guide = FALSE, values = c("black", "white")) +
              #scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
              labs(title = pg.title, 
                   subtitle="(b) Extreme switches in female preferences",
                   y = "Strength of viability selection (s)", 
                   x = expression(paste("Strength of sexual selection (", alpha, ")"))) +
              scale_fill_gradientn(colours = c("white", pal_female(100)), 
                                   limits = c(0,20), name = "Replicates with extreme circuit present",
                                   guide = guide_colorbar(frame.colour = "black", ticks.colour = "black")) +
              geom_rect(data=cust.highlight.cells, size=1, fill=NA, colour="black",
                  aes(xmin=as.numeric(a.ajusted) - 0.5, xmax=as.numeric(a.ajusted) + 0.5, ymin=as.numeric(s.ajusted) - 0.5, ymax=as.numeric(s.ajusted) + 0.5)) +
       theme(panel.background = element_blank(),
                          axis.line = element_line(colour = "black"),
                          axis.text=element_text(size=8),
                          #strip.text.x = element_blank(),
                          panel.grid.major = element_blank(), 
                          panel.grid.minor = element_blank(),
                          panel.border = element_rect(colour = "black", fill=NA, size=0.5),
                          plot.title = element_text(size=10),
                          plot.subtitle = element_text(size=9),
                          legend.position="bottom", legend.direction="horizontal",
                          legend.text=element_text(size=8),
                          legend.title=element_text(size=8),
                          aspect.ratio = 1)
      
    #print(heat.pref) 
    
    #TRAIT HEATMAP (group size=30)
    heat.trait <- ggplot(data = trait.summ, aes(x=as.factor(a), y=as.factor(s), fill=ta.nruns.variation.maintained_fuzzy)) + 
              geom_tile(colour="white",size=0.25) +
              # cell labels
              geom_text(aes(label = round(ta.first.stabilized.gen.fuzzy_median, 1), 
                                    color = ta.nruns.variation.maintained_fuzzy>15), size=2.5) +
              scale_color_manual(guide = FALSE, values = c("black", "white")) +
              #scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
              labs(title = pg.title, 
                   subtitle="(c) Maintenance of genetic variation in male trait",
                   y = "Strength of viability selection (s)", 
                   x = expression(paste("Strength of sexual selection (", alpha, ")")))  +
              scale_fill_gradientn(colours = c("white", pal_male(100)), 
                                   limits = c(0,20), name = "Replicates retaining trait variability", 
                                   guide = guide_colorbar(frame.colour = "black", ticks.colour = "black")) +
              geom_rect(data=cust.highlight.cells, size=1, fill=NA, colour="black",
                  aes(xmin=as.numeric(a.ajusted) - 0.5, xmax=as.numeric(a.ajusted) + 0.5, ymin=as.numeric(s.ajusted) - 0.5, ymax=as.numeric(s.ajusted) + 0.5)) +
       theme(panel.background = element_blank(),
                          axis.line = element_line(colour = "black"),
                          axis.text=element_text(size=8),
                          #strip.text.x = element_blank(),
                          panel.grid.major = element_blank(), 
                          panel.grid.minor = element_blank(),
                          panel.border = element_rect(colour = "black", fill=NA, size=0.5),
                          plot.title = element_text(size=10),
                          plot.subtitle = element_text(size=9),
                          legend.position="bottom", legend.direction="horizontal",
                          legend.text=element_text(size=8),
                          legend.title=element_text(size=8),
                          aspect.ratio = 1)
      
    #print(heat.trait) 
    
#pdf(paste("Fig_compiled_fd.fa", "pdf", sep="."), width=16, height=7, onefile = TRUE)
gridExtra::grid.arrange(ex.ts.plots, heat.pref, heat.trait, nrow = 1, ncol=3)  

}   
    
    
    
#end plotting (outside loop so that all plots are in single file)
#dev.off()


```

